Benchmarking neuromorphic systems with Nengo

### 2015-09-14 paper/rebuttals.txt

We have verified that the error bars reported are bootstrapped 95%
confidence intervals, as claimed in the figure captions of
Figures 5 and 6. As previously mentioned, we also replicated
the paradoxical slowdown with probes removed; however,
with the removal of Figure 7, we believe that additional text attesting
to the validity of the confidence intervals is not warranted.

> 3. In Figure 3, the meaning of negative frequencies (-1 Hz, -2 Hz)
> should be at least briefly mentioned. Presumably, since this is a
> circular oscillator, it is being run in reverse. However, it's not
> clear to me why negative frequencies should be at all different than
> positive frequencies, and how the range -2 Hz to 2 Hz was chosen. A
> citation that references this range in biology for CPGs, for example,
> would bolster this choice.

The meaning of negative frequencies has been made explicit in text;
line 239 now notes that "negative frequencies indicate
oscillations in the reverse direction." In that line, we have
also added a footnote with a citation to Duysens and Van de Crommert, 1998,
which provides some evidence that locomotive CPGs may run in reverse
for backward locomotion.

In terms of why the range -2 to 2 was chosen,
this was because we have used oscillators of these frequencies
in previous models. In line 219 we note that
"We test each model with parameters typically used
in large-scale models, but make these parameterized models
available ... for those who wish to explore additional cases."
The parameter is trivial to change, up to an upper frequency limit
which depends on the properties of the neurons in use;
we would be happy to show results for a different set of frequencies
if that would be preferable.

@@

5. Other comments on the results.
---------------------------------

> I found the description of the Figures on accuracy too short. Even
> if the accuracy results themselves are not the main focus of the
> article, some more explanations for the results should be given. In
> particular, in Figs. 1 and 2, why does the Nengo reference backend
> show a significantly higher variability than the Brainstorm and
> distilled backends? In Figs. 3 and 4, where do the deviations in the
> SpiNNaker backend come from (my first guess would be the employed
> 1ms time step, which is larger than the 0.1ms typically used in
> software simulators)?

We have added additional details in the section 4.2 on accuracy.
These additional details are listed below.

Briefly, we do not have a satisfying explanation
of why the reference backend shows higher variability than
the Brainstorm and distilled backends.
We discuss this briefly in the additional details below
and have added a sentence to the discussion
indicating that we will examine this result in the future
(lines 398-399: "We also plan to investigate
why the reference and OpenCL backends
have higher variability than other backends on
the simple feedforward models.")

@@

We are still in the progress of investigating the deviations
in the SpiNNaker backend; however, we are collaborating with
the SpiNNaker group to improve accuracy on these examples.
The 1ms time step is not the issue here, however,
as all of the backends tested use a 1ms time step.

Figure 1:

  The Brainstorm and distilled backends have the least variability and
  the SpiNNaker backend has the most variability, though the median
  RSME is the same across all backends.  One driver of these
  differences is in how each backend handles ensembles operating near
  the edge of their representational range (i.e., the radius of the
  ensemble).  The outlier with highest RMSE represents a model
  instance in which the static vector target was at the extreme of the
  representational range; the SpiNNaker backend's relatively high RMSE
  on this example indicates that it may not perform as well as other
  backends in this situation.  The SpiNNaker backend uses signed fixed
  point numbers with 16 digits before the decimal point, and 15 digits
  after the decimal point; accuracy could be improved by optimizing
  the backend's internal calculations to use as many digits as
  possible.  For all backends, accuracy could be improved by
  increasing the representational range and increasing the number of
  evaluation points generated when solving for decoding weights.

@@

Figure 2:

  Again, it is likely that performance at the extremes of the
  representational range is responsible for the SpiNNaker backend's
  reduced accuracy, though we have not investigated this in detail.
  This benchmark also suggests that the reference and OpenCL backends
  have more variability than other backends, particularly the
  Brainstorm and distilled backends.

@@

Figure 3:

  In this example, differences in how each backend implements synaptic
  filtering are more pronounced than in previous examples, as the
  recurrent connection is responsible for the dynamics of the
  oscillation.  Relatively large variability in the reference, OpenCL
  and SpiNNaker backends suggests that their synaptic filtering
  implementations should be examined for potential accuracy
  improvements.

### 2015-09-14 paper/paper.tex
Incorporate feedback from Terry

Figure~\ref{fig:01} gives accuracy results
for each backend on the chained communication channel model.
The boxplot shows that all five backends can
implement this model accurately,
despite five layers of processing
that each introduce noise.
The Brainstorm and distilled backends
have the least variability
and the SpiNNaker backend has the most variability,
though the median RSME is the same across all backends.
One driver of these differences
is in how each backend implement synaptic filtering.
The reference and OpenCL backends,
for example,
implement arbitrary linear time-invariant filters,
of which the lowpass filter used in this model
is a specific example.
The Brianstorm and distilled backends
only support lowpass filters,
and therefore implement it
more accurately.
Another reason for these differences
is in how each backend handles
ensembles operating near the edge
of their representational range
(i.e., the radius of the ensemble).
The outlier with highest RMSE represents
a model instance in which the static vector target
was at the extreme of the representational range;
the SpiNNaker backend's relatively high RMSE
on this example indicates that it
may not perform as well as other backends
in this situation.
However, accuracy could be improved
by increasing the representational range,
though doing so may necessitate an increase
in the number of evaluation points generated
when solving for decoding weights.

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
is less accurate than the other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.
Again, the more generic synaptic filtering model
used in the reference and OpenCL backends
is the likely cause of increased variability
on this model.
It is also likely that performance at the extremes
of the representational range
is responsible for the SpiNNaker backend's
reduced accuracy,
though we have not investigated this in detail.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
The median accuracy of the SpiNNaker is
the same as the distilled and Brainstorm backends,
but there are a high proportion of model instances
with less accuracy than the median,
as compared to other backends.
In this example, differences in synaptic filtering
are more pronounced than in previous examples,
as the recurrent connection is responsible
for the dynamics of the oscillation.
This result indicates that the synaptic filtering model
of SpiNNaker may also warrant closer examination
for accuracy.

@@

Differences in accuracy benchmarks
are also important for individual backends.
The improved accuracy
for the SpiNNaker backend
when passthrough nodes are removed
from the basal ganglia model
has resulted in ongoing work to
automatically remove passthrough nodes
in the build phase of the SpiNNaker backend.
Similarly, we will investigate
instances of the product model
in which the SpiNNaker backend
is less accurate than other backends,
though we suspect that its
lack of floating point hardware
may be responsible.

>>>>

Figure~\ref{fig:01} gives accuracy results
for each backend on the chained communication channel model.
The boxplot shows that all five backends can
implement this model accurately,
despite five layers of processing
that each introduce noise.
The Brainstorm and distilled backends
have the least variability
and the SpiNNaker backend has the most variability,
though the median RSME is the same across all backends.
One driver of these differences
is in how each backend handles
ensembles operating near the edge
of their representational range
(i.e., the radius of the ensemble).
The outlier with highest RMSE represents
a model instance in which the static vector target
was at the extreme of the representational range;
the SpiNNaker backend's relatively high RMSE
on this example indicates that it
may not perform as well as other backends
in this situation.
The SpiNNaker backend uses
signed fixed point numbers
with 16 digits before the decimal point,
and 15 digits after the decimal point;
accuracy could be improved
by optimizing the backend's
internal calculations to use as
many digits as possible.
For all backends, accuracy could be improved
by increasing the representational range
and increasing the number of evaluation points generated
when solving for decoding weights.

Figure~\ref{fig:02} gives accuracy results
for the two-dimensional product model.
In this case, while all of the backends
perform well, the SpiNNaker backend
is less accurate than the other four backends.
Its performance, however, is still well within
acceptable ranges for the two-dimensional product model.
Again, it is likely that performance at the extremes
of the representational range
is responsible for the SpiNNaker backend's
reduced accuracy,
though we have not investigated this in detail.
This benchmark also suggests that
the reference and OpenCL backends
have more variability than other backends,
particularly the Brainstorm and distilled backends.

Figure~\ref{fig:03} gives accuracy results
for the controlled oscillator model.
The reference and OpenCL backends perform well here,
as do the distilled and Brainstorm backends.
The median accuracy of the SpiNNaker is
the same as the distilled and Brainstorm backends,
but there are a high proportion of model instances
with less accuracy than the median,
as compared to other backends.
In this example, differences in how each backend
implements synaptic filtering
are more pronounced than in previous examples,
as the recurrent connection is responsible
for the dynamics of the oscillation.
Relatively large variability in the
reference, OpenCL and SpiNNaker backends
suggests that their synaptic filtering implementations
should be examined for potential accuracy improvements.

@@

Differences in accuracy benchmarks
are also important for individual backends.
The improved accuracy
for the SpiNNaker backend
when passthrough nodes are removed
from the basal ganglia model
has resulted in ongoing work to
automatically remove passthrough nodes
in the build phase of the SpiNNaker backend.
Similarly, we will investigate
instances of the product model
in which the SpiNNaker backend
is less accurate than other backends,
though we suspect that its
lack of floating point hardware
may be responsible.
We also plan to investigate
why the reference and OpenCL backends
have higher variability than other backends on
the simple feedforward models.
